---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggthemes)
library(gridExtra)
```

Learning a bit of R to take advantage of `ggplot`, since it looks neat.

The first thing we can do with Shiva's claims is that 

```{r}
polygon <- data.frame(
  x = c(0, 100, 100, 0),
  y = c(100, 0, -100, 0)
)

# Simulate drawing from the delta(X) distribution
numPoints <- 100
linspace <- seq(1,100,length=numPoints)
scatter <- data.frame(
  x = linspace,
  y = -1*linspace
)
zeroes <- geom_point(data=scatter, mapping=aes(x=x, y=y))

# Draw from the uniform distribution [0, 100]
scatter <- data.frame(
  x = linspace,
  y = -1*linspace + runif(numPoints, min=0, max=100)
)
uniform <- geom_point(data=scatter, mapping=aes(x=x, y=y))

# Draw from a linear model with log-normally distributed noise and a slope less than 1
lm <- function(slope, intercept, sd, numPoints, addedLine) {
  lm_ <- slope*linspace + intercept + rnorm(numPoints, mean=0, sd=sd)
  
  # reject samples to avoid vals that exceed [addedLine,100 + addedLine]
  while (any(lm_+addedLine < addedLine)) {
    idx <- lm_+addedLine < addedLine
    lm_[idx] = slope*linspace[idx] + intercept + rnorm(length(lm_[idx]), mean=0, sd=sd)
  }
  while (any(lm_+addedLine > 100+addedLine)) {
    idx <- lm_+addedLine > 100+addedLine
    lm_[idx] = slope*linspace[idx] + intercept + rnorm(length(lm_[idx]), mean=0, sd=sd)
  }

  return(lm_)
}

scatter <- data.frame(
  x = linspace,
  y = -1*linspace + lm(0.5, 0, 20, numPoints, -1*linspace)
)
linear1 <- geom_point(data=scatter, mapping=aes(x=x, y=y))
    
scatter <- data.frame(
  x = linspace,
  y = -1*linspace + lm(1.5, 0, 20, numPoints, -1*linspace)
)
linear2 <- geom_point(data=scatter, mapping=aes(x=x, y=y))


# Draw from a linear model with log-normally distributed noise and a slope more than 1

p <- ggplot(polygon, aes(x = x, y = y)) + geom_polygon(alpha=0.15)

p1 <- p + zeroes + theme_light()
p2 <- p + uniform + theme_light()
p3 <- p + linear1 + theme_light()
p4 <- p + linear2 + theme_light()

grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

```{r}

```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

```{r}

```

